Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-09 arxiv-version/main.tex
literature

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We identify two notions of linear separability:
strong linear separability and weak linear separability.
\begin{enumerate}
\item Under the strong linear separability condition, we design an efficient
algorithm that achieves a near-optimal mistake bound of $O\left(\frac{K}{\gamma^2} \right)$.

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a reduction from weak linear separability to strong linear separability
via an infinite dimensional monomial feature map, which in turn is inspired by
\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}
\end{abstract}

>>>>

\begin{abstract}
We study the problem of efficient online multiclass linear classification with
bandit feedback, where all examples belong to one of $K$ classes and lie in the
$d$-dimensional Euclidean space. Previous works have left open the challenge of
designing efficient algorithms with finite mistake bounds when the data is
linearly separable by a margin $\gamma$. In this work, we take a first step
towards this problem. We consider two notions of linear separability:
strong linear separability and weak linear separability.

\begin{enumerate}
\item Under the strong linear separability condition, we design an efficient
algorithm that achieves a near-optimal mistake bound of $O\left(\frac{K}{\gamma^2} \right)$.

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak
linear separable examples to strong linear separable examples, which in turn is inspired by
\cite{Klivans-Servedio-2008}'s work on learning intersection of halfspaces.
\end{enumerate}
\end{abstract}

### 2019-01-09 arxiv-version/notions-of-linear-separability.tex
literature

\begin{definition}[Linear separability]
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, let $K$ be a positive
integer, and let $\gamma$ be a positive real number. Let $(x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T)$ be labeled examples in $V \times \{1,2,\dots,K\}$.

The examples are said to be \emph{linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1 \; , \\
\label{equation:linear-separability-2}
\forall t \in \{1,2,\dots,T\} \quad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \qquad \qquad \ip{x_t}{w_{y_t}} & \ge \ip{x_t}{w_i} + \gamma \; .
\end{align}

The examples are said to be \emph{strongly linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:strong-linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1 \; , \\
\label{equation:strong-linear-separability-2}
\forall t \in \{1,2,\dots,T\} \qquad \qquad \ip{x_t}{w_{y_t}} &\ge \gamma \; , \\
\label{equation:strong-linear-separability-3}
\forall t \in \{1,2,\dots,T\} \qquad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \qquad \qquad \ip{x_t}{w_i} & \le - \gamma \; .
\end{align}
\end{definition}

The notion of linear separability with a margin is standard. It used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. For completeness, we state the
algorithm in Appendix~\ref{section:multiclass-perceptron-proofs}. It is a
folklore result that if a set of labeled examples is separable with a margin
$\gamma$ and the norm of the examples is bounded by $R$ then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. Another folklore result is that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any deterministic algorithm must make
$\left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the worst case. For
completeness, we give proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs}.

The notion of strong linear separability is new.

@@

>>>>

\begin{definition}[Linear separability]
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $(x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T)$ be labeled examples in $V \times \{1,2,\dots,K\}$.

The examples are said to be \emph{linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1 \; , \\
\label{equation:linear-separability-2}
\forall t \in \{1,2,\dots,T\} \quad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \qquad \qquad \ip{x_t}{w_{y_t}} & \ge \ip{x_t}{w_i} + \gamma \; .
\end{align}
\label{def:linear-sep}
\end{definition}

\begin{definition}[Strong Linear separability]
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number. Let $(x_1, y_1), (x_2,
y_2), \dots, (x_T, y_T)$ be labeled examples in $V \times \{1,2,\dots,K\}$.

The examples are said to be \emph{strongly linearly separable with a
margin $\gamma$} if there exist vectors $w_1, w_2, \dots, w_K \in V$ such
that
\begin{align}
\label{equation:strong-linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1 \; , \\
\label{equation:strong-linear-separability-2}
\forall t \in \{1,2,\dots,T\} \qquad \qquad \ip{x_t}{w_{y_t}} &\ge \gamma \; , \\
\label{equation:strong-linear-separability-3}
\forall t \in \{1,2,\dots,T\} \qquad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \qquad \qquad \ip{x_t}{w_i} & \le - \gamma \; .
\end{align}
\label{def:strong-linear-sep}
\end{definition}

The notion of linear separability with a margin is standard. It is used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. For completeness, we state the
algorithm in Appendix~\ref{section:multiclass-perceptron-proofs}. It is a
folklore result that if a set of labeled examples is separable with a margin
$\gamma$ and the norm of the examples is bounded by $R$ then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. Another folklore result is that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any deterministic algorithm must make
$\left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the worst case. For
completeness, we give proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs}.

The notion of strong linear separability is new.
\chicheng{I think the notion has appeared in~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}, however,
they proposed an algorithm that doesn't achieve a finite mistake bound in the bandit
setting under this assumption (although they claimed that they have done so, the proof was
wrong..)}

@@

\chicheng{Shall we change the $\gamma$ in Equations \eqref{equation:strong-linear-separability-2}, \eqref{equation:strong-linear-separability-3} to $\gamma/2$, to simplify the presentation?}

### 2019-01-09 arxiv-version/related-work.tex
literature

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the $\gamma$-weak
linearly separable setting are proposed: one with a mistake bound of $O(k^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{k^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret in the general setting, and has a $O(\sqrt{T})$ mistake
bound in the $\gamma$-weak linearly separable setting. The polynomial dependence
of \textsc{Banditron}'s mistake bound on the time horizon is undesirable for
problems with a long time horizon. One key open question left
by~\cite{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can deesign
computationally efficient algorithms that achieve mistake bounds that match or
improve over those of inefficient algorithms. In this paper, we take a step
towards answering this question, showing that efficient algorithms with mistake
bounds quasipolynomial in the margin parameter can be obtained.

@@

Many works consider bandit multiclass classification in the general
non-separable setting. \cite{Abernethy-Rakhlin-2009} poses the open problem of
whether one can design and efficient algorithm to get a $O(\sqrt{T})$ regret
against some reasonable loss functions. \cite{Crammer-Gentile-2013} show that
such an algorithm can be obtained, under the assumption that the distribution of
the label $y_t$ conditioned on feature vector $x_t$ satisfies certain parametric
noise assumption. \cite{Hazan-Kale-2011} developed the \textsc{Newtron}
algorithm, which has a regret between $O(\ln T)$ and $O(T^{2/3})$ against the
multiclass logistic loss, where the exact order of the regret depends on the
diameter of the benchmark class. In particular, if the diameter is $O(\ln T)$,
its regret bound would become $O(T^{2/3})$. The SOBA algorithm by
\cite{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
$\widetilde{O}(\sqrt{T})$ against the
$\eta$-loss~\cite{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, their regret
bound does not depend sensitively on the diameter of the benchmark class.
\cite{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
it doubly-exponentially improves over \cite{Hazan-Kale-2011}'s regret on its
dependence on the diameter of the benchmark class. Recently,
\cite{Foster-Krishnamurthy-2018} developed a rich theory on contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\frac{1}{K} \sum_{i=1}^K \phi( f_i(x) )$, where $\calF$ contains function $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ and $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
for instance parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$.

@@

>>>>

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the standard
linearly separable setting are proposed: one with a mistake bound of $O(K^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret against the multiclass hinge loss
in the general setting, and has a $O(\sqrt{T})$ mistake
bound in the $\gamma$-weak linearly separable setting. The polynomial dependence
of \textsc{Banditron}'s mistake bound on the time horizon is undesirable for
problems with a long time horizon. One key open question left
by~\cite{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can design
computationally efficient algorithms that achieve mistake bounds that match or
improve over those of inefficient algorithms. In this paper, we take a step
towards answering this question, showing that efficient algorithms with mistake
bounds quasipolynomial in the margin parameter can be obtained.

@@

Many works consider bandit multiclass classification in the general
non-separable setting. \cite{Abernethy-Rakhlin-2009} poses the open problem of
whether one can design an efficient algorithm to get a $O(\sqrt{T})$ regret
against some reasonable loss functions. \cite{Crammer-Gentile-2013} show that
such an algorithm can be obtained, provided that the distribution of
the label $y_t$ conditioned on feature vector $x_t$ satisfies certain parametric
noise assumption. \cite{Hazan-Kale-2011} developed the \textsc{Newtron}
algorithm, which has a regret between $O(\ln T)$ and $O(T^{2/3})$ against the
multiclass logistic loss, where the exact order of the regret depends on the
diameter of the benchmark class. In particular, if the diameter is $O(\ln T)$,
its regret bound would become $O(T^{2/3})$. The \textsc{SOBA} algorithm by
\cite{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
$\widetilde{O}(\sqrt{T})$ against the
$\eta$-loss~\cite{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, its regret
bound does not depend sensitively on the diameter of the benchmark class.
\cite{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
it doubly-exponentially improves over \cite{Hazan-Kale-2011}'s regret on its
dependence on the diameter of the benchmark class.

@@

The online bandit multiclass learning is a special case of the contextual bandits
problem~\cite{Auer-2003, Langford-Zhang-2007}, where the cost is the classification error.
A series of works in the contextual bandits learning literature focuses on
oracle-efficiency, that is, they assume access to an
oracle that given a set of cost-sensitve learning examples,

Recently, \cite{Foster-Krishnamurthy-2018} developed a rich theory on contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\frac{1}{K} \sum_{i=1}^K l_t(i) \phi( f_i(x) )$, where $\calF$ contains function $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ and $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
for instance parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$.
